\chapter{DRAFT: Coarray Features}
\label{chap:DRAFT: Coarray Features}

%-------------------------------------------------------%
% ENVIRONMENT 
%-------------------------------------------------------%
\newcommand{\onlyF}{{\tt [F]}} 
\newcommand{\onlyC}{{\tt [C]}} 

\newenvironment{Constraints}{\subsubsection*{Constraints}
 \begin{enumerate}}{\end{enumerate}}
\newenvironment{Constraints F}{\subsubsection*{Constraints {\onlyF}}
 \begin{enumerate}}{\end{enumerate}}
\newenvironment{Constraints C}{\subsubsection*{Constraints {\onlyC}}
 \begin{enumerate}}{\end{enumerate}}

\newenvironment{Restriction}{\subsubsection*{Restriction}
 \begin{itemize}}{\end{itemize}}

\newcommand{\NEW}[1]{\mytextcolor{red}{#1}}

%-------------------------------------------------------%
% PREFACE
%-------------------------------------------------------%

\framebox[0.9\textwidth][c]{
\rule[-5mm]{0mm}{10mm}
This chapter is a proposal document to be added before 
Section~\ref{chap:Support for the Local-view Programming}.
}

\bigskip


%-- COARRAY AS THE LOCAL-VIEW
For the local-view programming, {\XMP} supports the coarray features
as a part of the language specifications.
%-- SPECIFICATION RANGE OF THE XMP COARRAY
{\XMPF} contains all coarray features defined in the standard Fortran~2008
(ISO/IEC 1539-1:2010) with few incompatibility described in 
Section~\ref{sec:Compatibility with the Fortran Standard}
and includes some important intrinsic procedures defined in 
the standard Fortran~2015.
Also {\XMPC} contains the coarray features which was designed based on
the ones of {\XMPF}.


%-------------------------------------------------------%
\section{Introduction for Coarrays}
\label{sec:Introduction for Coarrays}
%-------------------------------------------------------%

%=======================================
\subsubsection*{Image and image index}
%=======================================
The local-view programming model is a Single Program Multiple Data (SPMD) model.
Each replication of the program is called an {\bf image}.
Every image has a different {\bf image index}, which is an integer number 
between one and the number of images.
The number of images is not determined until the program execution.

In {\XMP}, a virtual array of the whole images with any number of dimensions 
is called an {\bf image array}. 
Each array element of an image array is corresponding to an image index
in the array element order of Fortran.
The extent of the final (outermost) dimension is not determined until
the program execution because it depends on the number of images.

The images are mapped one-by-one to the execution nodes.
The correspondence between images and nodes is defined later.
%in Chapter~\ref{chap:Interoperability of Global- and Local-views}.
Inquire functions about the image index and the number of images 
are described in Section~\ref{sec:Intrinsic Procedures}.


%=======================================
\subsubsection*{Coarray}
%=======================================
%-- what is coarray
A {\bf coarray} is an object that has a corresponding image array.
Unlike a usual object (non-coarray), a coarray is allowed to be
referred from other images. 
Each image has its own coarrays and can reference and define 
coarrays on all images each other.

%-- coshape
The shape of the image array corresponding to a coarray is 
called a {\bf coshape} of the coarray.
A coshape is specified with a {\it coarray-spec} in the declaration of the 
coarray variable or the coarray pointer (described later).
%-- corank, cobounds, coextend
The number of dimensions of a coshape is called a {\bf corank}.
For each dimension of a coshape, the lower and upper bounds and the extent
are called {\bf lower and upper cobounds} and {\bf coextent}, respectively.


%-- [F] category of coarray
{\onlyF} A {\it coarray-spec} can appear in a type declaration statement 
and in a component definition statement.
Entities declared with a {\it coarray-spec} are categorized into three:
\begin{itemize}
\item A {\bf static coarray} is a coarray that is not a dummy argument and is
non-allocatable. 
It must have the SAVE attribute explicitly or implicitly.
\item A {\bf dummy coarray} is a coarray that is a dummy argument and is 
non-allocatable.
The actual argument corresponding to a dummy coarray must be a static coarray,
a dummy coarray, an allocatable coarray, or a subobject of them.
\item An {\bf Allocatable coarray} is a coarray that is allocatable.
An ultimate component\footnote
{A component is an ultimate component of the structure if it is of 
a basic type or is allocatable or a pointer.
An ultimate component of a component is an ultimate component of the structure,
recursively.}
of a non-coarray structure can also be an allocatable coarray.
\end{itemize}

%-- [C] category of coarray
{\onlyC} A {\it coarray-spec} can appear in a {\it declaration} and 
in a {\it parameter-declaration}. 
Entities declared with a {\it coarray-spec} are categorized into three:
\begin{itemize}
\item A {\bf static coarray} is a coarray that is not a dummy argument and
is of a basic, structure or array type. If it is an array, the array 
element must be of a basic, structure or array type.
It must have the {\tt static} or {\tt extern} storage class.
\item A {\bf dummy coarray} is a coarray that is a dummy argument and
is of a basic, structure or array type. If it is an array, the array 
element must be of a basic, structure or array type.
The actual argument corresponding to a dummy coarray must be the name
of a static corray, dummy coarray, or a coarray pointer.
\item A {\bf coarray pointer} is a pointer to a coarray called a 
{\bf target coarray}.
An ultimate component of a non-coarray structure can also be a
coarray pointer.
\end{itemize}

%-- life span of coshape
A static coarray is allocated previously and is static during the program 
execution. 
The coshape of a static coarray is explicitly specified in the declaration of the 
variable.
%
An allocatable coarray is dynamically allocated and deallocated at
the ALLOCATE and DEALLOCATE statements. 
A coarray pointer is dynamically allocated and freed by the intrinsic 
functions. 
The coshapes of them are determined at the allocation time and are retained 
until the deallocation/freeing time without regard to the scoping units.
%
A dummy coarray is an allocated object. 
The coshape of a dummy coarray is explicitly re-specified in the declaration of the
variable. Even if the corresponding actual argument is an allocatable coarrray or
a coarray pointer, the specification of the coshape is valid during the execution
in the scope.
%


\subsubsection*{Coarray Container}

A non-coarray structure object can have ultimate (leaf) components as 
allocatable coarrays (in {\XMPF}) or as coarray pointers (in {\XMPC}).
The strucure object is called a {\bf coarray container} in {\XMP}.

{\onlyF} A coarray container must be a scalar, may not be a pointer
or an allocatable, may not be a coarray, and may not be a function result.

{\onlyC} A coarray container may not be a coarray and may not be a function
result.


\subsubsection*{Cosubscript}
A coarray on the different image can be referenced and defined
by referring the coarray with {\bf cosubscripts}, which
is an array element of an image array.
See Section~\ref{sec:Allocation and Deallocation of Coarrays} for the detail.



\input{ch-coarray-decl}

\input{ch-coarray-darg}


%-------------------------------------------------------%
\section{Memory Allocation of Coarrays}
\label{sec:Memory Allocation of Coarrays}
%-------------------------------------------------------%

%============================================
\subsection{[F] Allocation of allocatable coarray}
\label{sec:Allocation of allocatable coarray}
%============================================

TBD

%============================================
\subsection{[C] Allocation of coarray pointer}
\label{sec:Allocation of coarray pointer}
%============================================

A coarray pointer is a pointer to a coarray object that is
called a target coarray.
A target coarray is allocated and freed with library functions
{\tt xmp\_comalloc} and {\tt xmp\_cofree}, respectively.
A coarray pointer retains the address of an allocated target coarray.
To avoid aliasing between coarrays, a coarray pointer is not allowed to 
point to any named coarrays or its subobjects,
nor to point the same target coarray or its subobjects that is pointed from
the other coarray pointer.


\subsubsection*{Example}

The first line of the code fragment shown below 
declares coarray pointer {\bf y} pointing to an array coarray 
of double type. 
The second line allocates a target coarray of size 
{\tt sizeof(double)*10*20} with the first-dimension coextent 4.
The third line frees the coarray and the value of {\bf y} becomes
invalid.

\begin{center}
 \begin{minipage}{0.70\hsize}
  \begin{XCexampleR}
  double (*y)[10]:[][];
  y = xmp_comalloc(sizeof(double)*10*20, 4);
  xmp_cofree(y);
  \end{XCexampleR}
 \end{minipage}
\end{center}


%-------------------------------------------------------%
\section{Referrence and Definition to Remote Coarrays}
\label{sec:Referrence and Definition to Remote Coarrays}
%-------------------------------------------------------%

TBD
%\input{ch-coarray-accs}

%-------------------------------------------------------%
\section{Synchronization and Error Handling}
\label{sec:Synchronization and Error Handling}
%-------------------------------------------------------%

TBD

%-------------------------------------------------------%
\section{Intrinsic Procedures}
\label{sec:Intrinsic Procedures}
%-------------------------------------------------------%

TBD

% Intrinsic function {\tt xmp\_coarray\_malloc} allocates an allocatable 
% coarray, determines the coextents, and returns the address to be assigned 
% to the coarray pointer.


%-------------------------------------------------------%
\section{Compatibility with the Fortran Standard}
\label{sec:Compatibility with the Fortran Standard}
%-------------------------------------------------------%

TBD
