
\section{Glossary}
\subsection*{\Term{node}}
In a distributed memory system, a computation node, which may have several cores sharing main memory, has its own local memory. Each node is connected through a network. An \XMP program begins as a single thread of execution in each node.

\subsection*{\Term{node number}}
A unique number assigned to each of the nodes in the entire node set. The number starts from 1, larger than or equal to 1 and less than and equal to the number of nodes.Note that the mapping from the node number to the MPI rank is decided by the system. The image index of
the coarray mapping to the entire node set is equal to the node number. 

\subsection*{\Term{node set}}
A set of nodes

\subsection*{\Term{entire node set}, \Term{entire set of nodes}}
All nodes executing the program, or a set of these nodes. The entire node set is
decided when staring the program.

\subsection*{\Term{executing node set}, \Term{executing nodes}}
A node set executing a certain region of a program. The executing node
set that executes an entire program is the entire node set. The executing
node set of a task is the node set that executes the task. 

\subsection*{\Term{node array}}
A multi-dimensional array containing nodes. The node array has a
name and shape as it attributes.

\subsection*{\Term{executing node array}}
Node array that contains the executing node.

\subsection*{\Term{task}}
A specific instance of executable code and
its data environments executed in a set of nodes. In the context of
the program text, a set of statement executed by a set of nodes. A task
can be nested, and a nested task is executed as a subtask of an outer task.

\subsection*{\Term{template}}
A dummy array used to express an index space associated
with an array. Template is also used to describe the iteration space of a loop. A template has a name, a dimension, and an upper and  lower bound for each dimension as attributes.

\subsection*{\Term{replicated execution}}
Execution of the same code in different nodes. If the state at the starting point is the same and
the execution has only local side-effects, then the local state in each node remains the same. 

\subsection*{\Term{data mapping}}
The combination of the alignment and
distribution attributes used to describe how a data object is
allocated to nodes.

\subsection*{\Term{work mapping}}
Assignment of iterations to nodes in
a parallel loop and tasks to nodes.

\subsection*{\Term{image index}}
A number assigned to
each image of the coarray. The value is equal to or greater than 1. 
Note that the image index of the coarray mapping to the entire node set is
identical to the node number.

\subsection*{\Term{local}}
Execution of a program has
side-effects only on the data in the node. In this case, no communication
with other nodes occurs . 

\subsection*{\Term{non-local}}
Execution of a program requires
communication with other nodes and has side-effects with respect to other
nodes. 

\subsection*{\Term{global data}}
Data declared as a distributed array and shared by nodes.

\subsection*{\Term{local data}}
Data is allocated in each node and is referenced only
within the node.* collective  An operation must be executed by every
node in the executing node set in order to perform an operation together.

\subsection*{\Term{distribution}}
The partition of the index space of a data
object among a set of nodes according to a given pattern. The
{\tt distribute} directive is used to map the elements of a template onto a set of nodes. 

\subsection*{\Term{alignment}}
An attribute of a data object that establishes
the relationship between data objects for distribution. The {\tt align}
directive is used to describe the correspondence of the element of the data and the template. 

\subsection*{\Term{shadow}}
A data area is used to keep neighbor elements
temporarily in a distributed array. Shadow is an attribute of
a distributed array that is declared by the {\tt shadow} directive and is updated by the {\tt reflect} directive.

